---
title: "Temperature"
header-includes:
    - \usepackage{booktabs}
    - \usepackage{longtable}
    - \usepackage{array}
    - \usepackage{multirow}
    - \usepackage[table]{xcolor}
    - \usepackage{wrapfig}
    - \usepackage{float} \floatplacement{figure}{H}
    - \usepackage{colortbl}
    - \usepackage{pdflscape}
    - \usepackage{tabu}
    - \usepackage{threeparttable}
    - \usepackage[normalem]{ulem}
    - \usepackage{setspace} \doublespacing
    - \newcommand{\beginsupplement}{\setcounter{table}{0} 
    - \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}
    - \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  bookdown::pdf_document2:
    number_sections: TRUE
    extra_dependencies: "subfig"
bibliography: '`r system.file("references.bib", package="weatheR")`'
vignette: >
  %\VignetteIndexEntry{Temperature}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
library(weatheR)
library(tidyverse)
library(ggsci)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 600,
  fig.width = 7,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%"
)
```

\newpage
# Growing degree-days

Growing degree day can be calculated in several ways:

- **Method A and B**: Linear
$$GDD = \frac{(T_{max} + T_{min})}{2} - T_{base}$$


Methods A and B differ in their way of handling daily minimum and maximum air 
temperatures. In **method A**, the comparison to base temperature occurs
after calculating the average temperature whereas in **method B**, the 
comparison to base temperature is made before calculating average air 
temperature  using the following rules [@mcmaster_growing_1997]:

- if $T_{max} < T_{base}$ then, $T_{max} = T_{base}$ 
- if $T_{min} < T_{base}$ then, $T_{min} = T_{base}$

**That's why, it's important to specify the method used to calculate GDD.**


- **Method C**: Triangle 

- **Method D**: Trapezoidal

```{r rate-of-development, fig.cap="Rate of develoment as function of temperature for different growing degree-days (GDD) methods."}
# Cardinal temperatures
T_base <- 5
T_opt <- 15
T_opt_1 <- 10
T_opt_2 <- 20
T_max <- 25

# Calculation
rate_of_development <- 
  tibble(.rows = 31) %>% 
  mutate(
    T_air_min = seq(0, 30, 1),
    T_air_max = seq(0, 30, 1),
    T_air_avg = seq(0, 30, 1),
    # Method A
    method_a =
      t_growing_degree_days(
        T_air_avg = T_air_avg,
        T_base = T_base,
        method = "A"
      )[["rate_of_development"]],
    # Method B
    method_b =
      t_growing_degree_days(
        T_air_min = T_air_min,
        T_air_max = T_air_max,
        T_base = T_base,
        method = "B"
      )[["rate_of_development"]],
    # Method C
    method_c =
      t_growing_degree_days(
        T_air_avg = T_air_avg,
        T_base = T_base,
        T_opt = T_opt,
        T_max = T_max,
        method = "C"
      )[["rate_of_development"]],
    # Method D
    method_d = 
      t_growing_degree_days(
        T_air_avg = T_air_avg,
        T_base = T_base,
        T_opt_1 = T_opt_1,
        T_opt_2 = T_opt_2,
        T_max = T_max,
        method = "D"
      )[["rate_of_development"]]
  ) %>% 
  pivot_longer(cols = c(
    "method_a",
    "method_b",
    "method_c",
    "method_d"),
    names_to = "variable",
    values_to = "value") %>% 
  mutate(
    variable = str_replace_all(string = variable,
                               pattern = "method_a",
                               replacement = "Linear - Small grains"),
    variable = str_replace_all(string = variable,
                               pattern = "method_b",
                               replacement = "Linear - Corn"),
    variable = str_replace_all(string = variable,
                               pattern = "method_c",
                               replacement = "Triangle"),
    variable = str_replace_all(string = variable,
                               pattern = "method_d",
                               replacement = "Trapezoidal")
  ) 

# Plot
ggplot(rate_of_development,
       aes(x = T_air_avg,
           y = value,
           color = variable)) +
  geom_point(size = 3) +
  geom_line(linewidth = 1.2) +
  theme_bw() +
  theme(text = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "right",
        legend.box = "vertical") +
  xlab("Average air temperature (°C)") +
  ylab("Rate of development (1/d)") +
  labs(color = "GDD methods")
```


```{r growing-degree-days, fig.cap="Cumulative growing degree-days (GDD) calculated with different methods."}
# Cardinal temperatures
T_base <- 5
T_opt <- 15
T_opt_1 <- 10
T_opt_2 <- 20
T_max <- 25

# Calculation
growing_degree_days <- 
  weather_blue_grass_airport %>% 
  filter(date > "2022-10-01",
         date < "2023-04-20") %>% 
  mutate(
    T_air_avg = (T_air_max + T_air_min)/2,
    # Method A
    method_a =
      cumsum(
        t_growing_degree_days(
          T_air_avg = T_air_avg,
          T_base = T_base,
          method = "A"
        )[["growing_degree_days"]]
      ),
    # Method B
    method_b = 
      cumsum(
        t_growing_degree_days(
          T_air_min = T_air_min,
          T_air_max = T_air_max,
          T_base = T_base,
          method = "B"
        )[["growing_degree_days"]]
      ),
    # Method C
    method_c =
      cumsum(
        t_growing_degree_days(
          T_air_avg = T_air_avg,
          T_base = T_base,
          T_opt = T_opt,
          T_max = T_max,
          method = "C"
        )[["growing_degree_days"]]
      ),
    # Method D
    method_d = 
      cumsum(
        t_growing_degree_days(
          T_air_avg = T_air_avg,
          T_base = T_base,
          T_opt_1 = T_opt_1,
          T_opt_2 = T_opt_2,
          T_max = T_max,
          method = "D"
        )[["growing_degree_days"]]
      )
  ) %>% 
  pivot_longer(cols = c("method_a",
                        "method_b",
                        "method_c",
                        "method_d"),
               names_to = "variable",
               values_to = "value") %>% 
  mutate(
    variable = str_replace_all(string = variable,
                               pattern = "method_a",
                               replacement = "Linear - Small grains"),
    variable = str_replace_all(string = variable,
                               pattern = "method_b",
                               replacement = "Linear - Corn"),
    variable = str_replace_all(string = variable,
                               pattern = "method_c",
                               replacement = "Triangle"),
    variable = str_replace_all(string = variable,
                               pattern = "method_d",
                               replacement = "Trapezoidal")
  )

# Plot
ggplot(growing_degree_days,
       aes(x = date,
           y = value,
           color = variable)) +
  geom_point() +
  theme_bw() +
  theme(text = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = "right",
        legend.box = "vertical") +
  xlab("Average air temperature (°C)") +
  ylab("Growing degree days (°C/d)") +
  labs(color = "GDD methods")
```

# References
